Hydrodynamics and rheology of adhesive vesicle doublets

### 2019-01-17 vesicle_code/docs/adhesion/adhesion.tex
some changes from yn

Adhesion between a vesicle and a substrate has been extensively studied,
primarily focused on analyzing the static equilibrium
shapes~\cite{Seifert1990_PRA, ShiFengGao2006_ActaMechSin,
LinFreund2007_IntJSolidsStructures, das2008adhesion, zhang2009phase}
than the transient adhesion process; exceptions
include~\cite{cantat1999lift, suk-sei2001, BlountMiksisDavis2013_PRSa}.
Adhesion between bilayer membranes was first experimentally observed by
Evans and Metcalfe~\cite{EvansMetcalfe1984_BJ}, who measured the
reduction in the free energy per unit area of membrane-membrane contact
formation due to the Van der Waals' attraction. For a membrane-membrane
separation distance of $\sim 3$ nm, they estimated that the Hamaker
constant is approximately $5.8\times 10^{-14}$ ergs, which corresponds
to an adhesion energy density around $1$ mJ/m$^2$. The ratio of adhesion
energy per unit area to the bending energy (bending rigidity of lipid
bilayer membranes $\sim 10^{-19}$ J) gives a measure of vesicle
deformation in the presence of
adhesion~\cite{RamachandranAndersonLealIsraelachvili2010_Langmuir}. In
this work we focus on regimes where the ratio is of order unity, between
the weak adhesion (ratio $\ll 1$) and strong adhesion (ratio $\gg 1$)
regimes.  In this regime the membrane deformation may increase the
``contact area'' between two vesicles and enhance the adhesion effects.
The equilibration of vesicle membranes under adhesion in these regimes
have been well studied and documented
(see~\cite{RamachandranAndersonLealIsraelachvili2010_Langmuir} and
references therein). However it is unclear how the adhesion couples to
the vesicle hydrodynamics in such regime.  Our work in this paper aims
to address this question with quantitative characterization in terms of
physical parameters.

@@

The exponents $(m,n)$ depend on the geometry of the two objects under
adhesion~\cite{Book_IntermolecularSurfaceForces}: $(m,n)=(4,2)$
corresponds to two flat, planar surfaces interacting with each under a
long-range attraction and a short-range repulsion. On the other hand,
$(m,n) = (12,6)$ corresponds to the L-J potential between two molecules.
The adhesion potential in  the integral of
equation~\eqref{eqn:adhesionForce} is between two small patches of lipid
bilayer membrane. Thus we argue that a reasonable choice for $(m,n)$ is
between $(m,n) = (4,2)$ and $(m,n) = (12,6)$.

@@

When the force on each vesicle is a function of separation distance $h$ of the form (with an integer $m\ge 2$)
\begin{align}
\label{eq:Fh}
F(h) \sim H \frac{\delta^m}{h^{m+1}}\left(\left(\frac{\delta}{h}\right)^2-1\right),
\end{align}
attraction is dominant at ``large" scales ($h > \delta$) while repulsion
is dominant at ``small" scales ($h<\delta$).  Integration of
equation~\eqref{eq:dhdt} with the above $F(h)$ gives the relationship
between $t$ and $h$.  In this case the relationship involves an integral
of a function of dimensionless $\bar{h} \equiv h/\delta$:
\begin{align*}
t\frac{K_a^{2/3}}{\mu R_0^{10/3}}& = \int^{h(t)}_{h(0)}-\frac{dh}{h^3
  F^{1/3}} = -\frac{1}{H^{1/3}\delta^{5/3}}\int^{\bar{h}(t)}_{\bar{h}(0)}\frac{d \bar{h}}{\bar{h}^{(8-m)/3}(\bar{h}^2-1)^{1/3}},
\end{align*}
\begin{align}
\label{eq:teq_scaling}
t\sim \frac{\mu R_0^{10/3}}{K_a^{2/3}}\frac{1}{H^{1/3}\delta^{5/3}}.
\end{align}
Equation~\eqref{eq:teq_scaling} tells us that both adhesion strength $H$
and the separation distance $\delta$ affect the time it takes for a pair
of vesicles to reach equilibrium under adhesion force $F$ in
equation~\eqref{eq:Fh}.  In particular, the equilibration time $t$ is
proportional to (1) the separation distance $\delta^{-5/3}$ (thus
$t\rightarrow \infty$ as $\delta\rightarrow 0$), and (2) the adhesion
strength $H^{-1/3}$ (as in constant forcing
case~\cite{RamachandranLeal2010_PoF}).  From the above analysis we also
expect that the scaling of $t$ with respect to $\delta$ depends on the
adhesion potential, while the scaling with respect to adhesion strength
is independent of the specific form of the potential.

>>>>

Adhesion between a vesicle and a substrate has been extensively studied,
primarily focused on analyzing the static equilibrium
shapes~\cite{Seifert1990_PRA, ShiFengGao2006_ActaMechSin,
LinFreund2007_IntJSolidsStructures, das2008adhesion, zhang2009phase}
than the transient adhesion process; exceptions
include~\cite{cantat1999lift, suk-sei2001, BlountMiksisDavis2013_PRSa}.
%
% The short-range adhesive potential of a lipid bilayer membrane
% inevitably involves the membrane change redistribution and the membrane potential
% 
%  in relation the screening of charges in the electric double layer (EDL) right next to the
% membrane \cite{Book_IntermolecularSurfaceForces,LiraDimova}.
% When two vesicles are in close vicinity of each other within the electric double layer,  
% In this work we do not consider the electrostatic of two lipid bilayer membranes.
% 
Adhesion between bilayer membranes was first experimentally observed by
Evans and Metcalfe~\cite{EvansMetcalfe1984_BJ}, who measured the
reduction in the free energy per unit area of membrane-membrane contact
formation due to the Van der Waals' attraction. For a membrane-membrane
separation distance of $\sim 3$ nm, they estimated that the Hamaker
constant is approximately $5.8\times 10^{-14}$ ergs, which corresponds
to an adhesion energy density around $1$ mJ/m$^2$. The ratio of adhesion
energy per unit area to the bending energy (bending rigidity of lipid
bilayer membranes $\sim 10^{-19}$ J) gives a measure of vesicle
deformation in the presence of
adhesion~\cite{RamachandranAndersonLealIsraelachvili2010_Langmuir}. In
this work we focus on regimes where the ratio is of order unity, between
the weak adhesion (ratio $\ll 1$) and strong adhesion (ratio $\gg 1$)
regimes.  In this regime the membrane deformation may increase the
``contact area'' between two vesicles and enhance the adhesion effects.
The equilibration of vesicle membranes under adhesion in these regimes
have been well studied and documented
(see~\cite{RamachandranAndersonLealIsraelachvili2010_Langmuir} and
references therein). However it is unclear how the adhesion couples to
the vesicle hydrodynamics in such regime.  Our work in this paper aims
to address this question with quantitative characterization in terms of
physical parameters.

@@

The exponents $(m,n)$ depend on the geometry of the two objects under
adhesion~\cite{Book_IntermolecularSurfaceForces}: $(m,n)=(4,2)$
corresponds to two flat, planar surfaces interacting with each under a
long-range attraction and a short-range repulsion. On the other hand,
$(m,n) = (12,6)$ corresponds to the L-J potential between two molecules.
In our case,  the adhesion potential in  the integral of
equation~\eqref{eqn:adhesionForce} is between two small patches of lipid
bilayer membrane (because the lipid molecules are coarse-grained in the continuum modeling).
Thus a reasonable choice for $(m,n)$ between two coarse-grained membrane patches
 would be between those for two planar surfaces and two point molecules.

 Large value of $m$ corresponds to a sharp increase in repulsion force as two objects are within the separation distance $\delta$.
 This poses great challenges for simulations as the numerical time-marching can be very stiff (i.e., very small time step) for large $m$. 
 The adaptive time-stepping BIE scheme makes it possible to simulate vesicle adhesive dynamics with specified numerical precision for reasonable computation time.
 We explore several combinations of $(m,n)$ in the simulations 
 of two vesicles forming a doublet in a quiescent flow, 
 and found very little difference in both the dynamic evolution and equilibrium configuration between $(m,n)=(8,6)$ and $(m,n)=(4,2)$.

 %
%
% Thus we argue that a reasonable choice for $(m,n)$ is
%between $(m,n) = (4,2)$ and $(m,n) = (12,6)$.

@@

When the force on each vesicle is a function of separation distance $h$ of the form (with an integer $m\ge 2$)
\begin{align}
\label{eq:Fh}
F(h) \sim H \frac{\delta^m}{h^{m+1}}\left(\left(\frac{\delta}{h}\right)^{m-n}-1\right),
\end{align}
attraction is dominant at ``large" scales ($h > \delta$) while repulsion
is dominant at ``small" scales ($h<\delta$).  Integration of
equation~\eqref{eq:dhdt} with the above $F(h)$ gives the relationship
between $t$ and $h$.  In this case the relationship involves an integral
of a function of dimensionless $\bar{h} \equiv h/\delta$:
\begin{align*}
t\frac{K_a^{2/3}}{\mu R_0^{10/3}}& = \int^{h(t)}_{h(0)}-\frac{dh}{h^3
  F^{1/3}} = -\frac{1}{H^{1/3}\delta^{5/3}}\int^{\bar{h}(t)}_{\bar{h}(0)}\frac{d \bar{h}}{\bar{h}^{(8-m)/3}(\bar{h}^{m-n}-1)^{1/3}},
\end{align*}
\begin{align}
\label{eq:teq_scaling}
t\sim \frac{\mu R_0^{10/3}}{K_a^{2/3}}\frac{1}{H^{1/3}\delta^{5/3}}.
\end{align}
Equation~\eqref{eq:teq_scaling} tells us that both adhesion strength $H$
and the separation distance $\delta$ affect the time it takes for a pair
of vesicles to reach equilibrium under adhesion force $F$ in
equation~\eqref{eq:Fh}.  In particular, the equilibration time $t$ is
proportional to (1) the separation distance $\delta^{-5/3}$ (thus
$t\rightarrow \infty$ as $\delta\rightarrow 0$), and (2) the adhesion
strength $H^{-1/3}$ (as in constant forcing
case~\cite{RamachandranLeal2010_PoF}).  From the above analysis we also
expect that the scaling of $t$ with respect to $\delta$ depends on the
adhesion potential, while the scaling with respect to adhesion strength
is independent of the specific form of the potential.
